---
title: "Henry Hub Spot Gas Prices"
author: "Group 8"
date: "06/24/2023"
output:
  pdf_document:
    toc: yes
    toc_depth: 1
  word_document:
    toc: yes
    toc_depth: '1'
subtitle: 'STAT 626: Time Series Analysis'
---

```{r, echo=FALSE,warning=FALSE, message=FALSE}
library(knitr)
library(xts)
library(astsa)
library(timeSeries)
library(forecast)
```


```{r}
setwd("C:/Users/14699/Desktop/STAT 626 Summer 2023")
gas_data = read.csv("Henry_Hub_Natural_Gas_Spot_Price_Monthly.csv")
# transpose of dataframe
trans_gas <- t(gas_data)
# converting the result to dataframe
trans_gas <- as.data.frame(trans_gas)
dollars = gas_data$Dollars
head(gas_data)
summary(gas_data)
plot(density(dollars))
tsplot(dollars)
abline(reg=lm(dollars~time(dollars)), col=4)
```

```{r}
# Detrend series using linear regression
fit = lm(dollars ~ time(dollars), na.action=NULL, data=gas_data)
tsplot(resid(fit), main="Detrended Gas Spot Prices")


# Differenced series 
tsplot(diff(dollars), main="Differenced Gas Spot Prices")
```

## Here we're going to discuss which ones the transformations we've done (failed or looks good enough) Which is Log, diff(), and diff(log()). What they look like, residuals, normality assumption, and potential stationarity. 

```{r}
# Log Transform
tsplot(log(dollars), main = "Log Transformed Gas Spot Prices")
acf2(log(dollars), plot=TRUE)
qqnorm(log(dollars)); qqline(log(dollars))
lag.plot(log(dollars), 24)
```

```{r}
# Log differenced series 
tsplot(diff(log(dollars)), main="Log Differenced Gas Spot Prices")
acf2(diff(log(dollars)), plot = TRUE)
lag.plot(diff(log(dollars)), 24)
qqnorm(diff(log(dollars)));qqline(diff(log(dollars)))
```

# Differencing
```{r}
tsplot(diff(dollars))
acf2(diff(dollars), plot=TRUE)
lag.plot(diff(dollars), 24)
qqnorm(diff(dollars)); qqline(diff(dollars))

```



```{r}
# Sample ACFs
acf1(resid(fit), 48) # Detrended
acf1(diff(dollars), 48) # Differenced

mean(diff(dollars)) # drift

lag.plot(dollars, 36)
```



```{r}
dollars <- ts(dollars)
acf1(diff(log(dollars)), 48) # Differenced
plot(resid(diff(log(dollars))))
```

```{r}

## acf2(dollars, 36) # Original data

```

```{r}
dollars <- ts(dollars)
acf2(log(dollars), 36) # Log

```

```{r}

acf2(diff(dollars), 36) # Differenced

```


### DIFF Log. Transformation Model we would like to do explain successfully. Included are the acfs. Have decided to go on this diff(log()) transformation of the data
```{r}
tsplot(diff(log(dollars)))
acf1(diff(log(dollars)), 60) # Log Differenced
acf2(diff(log(dollars)))
lag.plot(diff(log(dollars)), 36)

qqnorm(diff(log(dollars)));qqline(diff(log(dollars)))
## Not quite normal. Tails are quite a ways off which makes it seem not as normal
```


```{r}
auto.arima(diff(log(dollars)))

```


# Here we need to make an ideal sarima/ARMA by experimenting by deciding how many lags for each 
```{r}

sarima(diff(log(dollars)), p=4,d=0,q=4)
```

```{r}
# Log Transformation
tsplot(dollars)
tsplot(log(dollars))
qqnorm(dollars); qqline(dollars) # Not normal
qqnorm(log(dollars)); qqline(log(dollars)) # Still not normal
```


```{r}
# Check for multicollinearity between d and d^2
plot(diff(dollars), diff(dollars)^2)
cor(diff(dollars), diff(dollars)^2) # Collinear

dollars_adj = dollars - mean(dollars)
plot(dollars_adj, dollars_adj^2)
cor(dollars_adj, dollars_adj^2) # Collinear
```


```{r}
# linear regression - EXAMPLE 3.5 with multiple variables
fit = lm(dollars~time(dollars), na.action=NULL)
summary(fit) # Awful


# Difference
```


```{r}
#attempting lin reg with lagged using
fit2 = dynlm::dynlm(diff(log(dollars))~time(diff(log(dollars))), na.action=NULL)
summary(fit2)

# linear regression with Lagged Variables
fitl = dynlm::dynlm(dollars ~ L(dollars,5))
summary(fitl) # Awful
tsplot(resid(fitl))

### ALL SIGNS POINT TO AR CHAPTER 4 ###

```

## Corrected Data frame of Gas Prices

```{r}
setwd("C:/Users/14699/Desktop/STAT 626 Summer 2023")
gas_data2 = read.csv("Henry_Hub_Natural_Gas_Spot_Price.csv")
gas_data2$Price = gas_data2$Henry.Hub.Natural.Gas.Spot.Price.Dollars.per.Million.Btu
gas_data2$Price = as.numeric(gas_data2$Price)
# transpose of dataframe
trans_gas2 <- t(gas_data2)
# converting the result to dataframe
trans_gas2 <- as.data.frame(trans_gas2)
# calculating reverse of dataframe
rev_gasframe <- rev(trans_gas2)
# transpose of reverse dataframe
rev_gasframe <- t(rev_gasframe)
# converting the result to dataframe
new_gas <- as.data.frame(rev_gasframe)
## View(new_gas) ## Ideally Ordered in normal Ascending timeline
plot(new_gas$Price,col="blue", type = "l", ylab = "Dollars per Million BTU",
       xlab = "Months since Jan 1997", main = "Natural Gas Price Index by EIA")
price = new_gas$Price
abline(reg=lm(new_gas$Price~time(new_gas$Price)), col=4)
```




