---
title: "Henry Hub Spot Gas Prices"
author: "Group 8"
date: "sys.Date()"
output:
  pdf_document:
    toc: yes
    toc_depth: 1
  word_document:
    toc: yes
    toc_depth: '1'
subtitle: 'STAT 626: Time Series Analysis'
---

```{r packages, echo=FALSE,warning=FALSE, message=FALSE}
library(knitr)
library(xts)
library(astsa)
library(lubridate)
library(MASS)
library(timeSeries)
library(forecast)
```

```{r setup, include=FALSE}
# Set Working Directory
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE) # Hide code
# knitr::opts_knit$set(root.dir = "G:/My Drive/STAT/STAT 626/Group Project")
getwd()
```


## Data Set Up 

```{r ReadData}
gas_data = read.csv("Henry_Hub_Natural_Gas_Spot_Price_Monthly.csv")
# transpose of dataframe
trans_gas <- t(gas_data)
# converting the result to dataframe
trans_gas <- as.data.frame(trans_gas)
# reorder so 1997 is at the beginning
dollars = ts(gas_data$Dollars)
gas_data = gas_data[order(time(dollars),decreasing = T),]
head(gas_data)
summary(gas_data)
plot(density(dollars))
tsplot(dollars)

```

### Detrending Data (Not using)

First attempt at a transformation using detrending. Detrending using linear regression to estimate a best fit line of the data. This line equation is then subtracted from the data to de-trend it. Then the data should look generally stationary. Our data doesn't sit as stationary around a trend thus this method does not work. 

```{r Detrend}
# Detrend series using linear regression
fit = lm(dollars ~ time(dollars), na.action=NULL, data=gas_data) # trend only model
tsplot(resid(fit), main="Detrended Gas Spot Prices")

summary(fit)
# intercept is 4.69
# B1 is -0.02899
# mut = 4.69 - 0.00298t
fitValues = predict(fit,newdata=time(dollars))

# fit abline over tsplot 
gas<-ts(gas_data$Dollars,start = c(1997,1),freq=12)

plot(gas, 
     type="l", 
     lwd=2, 
     col="blue" ,
     main = "Log and Differenced Natural Gas Price Index by EIA",
     ylab= "Dollars",xlab = "Year",
     xlim=c(1997,2023))
lines(fitValues)

par(mrow=c(1,2))
tsplot(dollars,col = 2,main="Natual Gas Price Index by EIA")
tsplot(dollars- fitValues,col=3,main="Detrended Natural Gas Price Index by EIA")
```

# Differencing Data Plot

Differencing with an order of 1 subtracts the last observation (t-1) from the current (t) as the new obseration value at time t. The length dataset looses one value. We had 317 and how have 316. 

```{r DifferencePlot}
tsplot(diff(dollars))
```

## Log transformation Plot

Log transformation reduces heteroskedasticity and extreme values

```{r LogTransform}
# mean and variance are much different after April 2009. Points to heteroskedasticity
mean(dollars[1:146]) ; var(dollars[1:146])
mean(dollars[147:length(dollars)]) ; var(dollars[147:length(dollars)])
tsplot(log(dollars))


par(mfrow=c(1,2))
qqnorm(dollars);qqline(dollars)
qqnorm(log(dollars),main="Log Normal Q-Q Plot");qqline(log(dollars))
```


## Differncing the log transformed data Plot

```{r LogDiffTransform}
tsplot(diff(log(dollars)),main="Log and Differenced Natural Gas Price Index",ylab="Differenced Log(Dollars)") # this reduces the effect of the extreme values observed at a couple times throughout the data.


gas<-ts(diff(log(gas_data$Dollars)),start = c(1997,1),freq=12)
plot(gas, type="l", lwd=2, col="blue" ,main = "Log and Differenced Natural Gas Price Index by EIA",
     ylab= "Differenced Log(Dollars)",xlab = "Year",
     xlim=c(1997,2023))
```


## ACF and PACF Plots 

### Original Dataset 
```{r OriginalACF}
acf2(dollars, 36) # Original data
```

### Differenced Dataset 

```{r DifferenceACF}
acf2(diff(dollars), 36) # Log
# The pact plot points to a AR(1) time series of the log transformation
```

### Log transformed Dataset

```{r LogACF}
acf2(log(dollars), 36) # Log
```

### Log and Differenced Time Series 

```{r CombinationACF}
acf2(diff(log(dollars)), 36,main="Series: Log and Differenced Natural Gas Index")
qqnorm(diff(log(dollars))); qqline(diff(log(dollars)))
```

## Model Estimation

### Auto ARMA package

```{r}
mod = auto.arima(diff(log(dollars)))
summary(mod)
plot(forecast(mod,h=5))
```
# Transformed Data with Yule-Walker, ols and mean, Sarima function and forecasting, With residuals
```{r}
data = diff(log(dollars))
d.yw = ar.yw(data, order=1)
exp(d.yw$x.mean)
d.yw$ar # mean estimate
exp(d.yw$var.pred) # error variance estimate
names(d.yw)

sarima(data, p=4, d=0, q=4)
sarima.for(data, n.ahead=12, p=4,d=0,q=4)
```
# SARIMA using P,D,Q parameters
```{r}
sarima(data, p=1,d=0,q=1, P=1, D=0, Q=1)

```

```{r}
ols = ar.ols(data, order.max=2, demean=FALSE, intercept=TRUE)
ols$aic

level = as.numeric(data[-1])
lags = as.numeric(data[-length(data)])

d.lm = lm(level ~ lags)

summary(d.lm)

```




```{r}
# Sample ACFs
acf1(resid(fit), 48) # Detrended
acf1(diff(dollars), 48) # Differenced

mean(diff(dollars)) # drift

lag.plot(dollars, 36)
```









```{r}
# Check for multicollinearity between d and d^2
plot(diff(dollars), diff(dollars)^2)
cor(diff(dollars), diff(dollars)^2) # Collinear

dollars_adj = dollars - mean(dollars)
plot(dollars_adj, dollars_adj^2)
cor(dollars_adj, dollars_adj^2) # Collinear
```


```{r}
# linear regression - EXAMPLE 3.5 with multiple variables
fit = lm(dollars~time(dollars), na.action=NULL)
summary(fit) # Awful
plot(fit)
```


```{r}
# linear regression with Lagged Variables
fitl = dynlm::dynlm(dollars ~ L(dollars,5))
summary(fitl) # Awful
tsplot(resid(fitl))

### ALL SIGNS POINT TO AR CHAPTER 4 ###
```


